---
title: "Process Energy and Emissions File"
---

I have an Excel file with data compiled from a few different sources about energy use and emissions for world countries.

The name of the file is `Energy and Emissions.xlsx` and it contains two sheets, as follows:

-   Energy_Emissions: the data
-   Source: notes about the source and a data dictionary

To make it easy to share this dataset I will read it, convert it to an `.RData` file, and then document it so that it can be bundled with package `packr`.

Load the packages needed to preprocess the data:

```{r}
library(tidyverse)
library(readxl)
library(countrycode)
```

Read the Excel file as a dataframe:

```{r}
RAW_DATA_metabolomic_of_saliva <- read_excel("RAW DATA metabolomic of saliva.xlsx")
```

The dataframe consists of 188 observations of 10 variables. Get a summary of the dataframe:

```{r}
summary(RAW_DATA_metabolomic_of_saliva)
```

The names of the samples are character, so will change to factor:

```{r}
RAW_DATA_metabolomic_of_saliva <- RAW_DATA_metabolomic_of_saliva %>%
  mutate(SAMPLES = factor(SAMPLES), GROUPS = factor(GROUPS))
```

New summary of the dataframe:

```{r}
summary(RAW_DATA_metabolomic_of_saliva)
```


The sources of these data are as follows:

-   DMP Article
(<https://www.sciencedirect.com/science/article/pii/S2352340920315560>) Accessed: November 6, 2023

-   Excel file
(<https://data.mendeley.com/datasets/wys5xd2vfg/1>) Accessed: November 6, 2023


Save to data folder:

```{r}
usethis::use_data(RAW_DATA_metabolomic_of_saliva,
                  overwrite = TRUE)
```
