---
title: "package1-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{package1-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r install-data-package, include=FALSE}
if (!require("package1", character.only = TRUE)) {
      remotes::install_github("gillec1/package1",
                        build_vignettes = TRUE)
  }
```

```{r setup}
library(package1)
```

#Defining the parameters within the excel file
- Groups: 53 women total, aged 21 to 68 years
  - Control = 27 healthy volunteers
  - EG-pre = 26 women with muscular temporomandibular disorder (TMD), whose saliva was taken before undergoing            treatment. Treatment involved applying hot packs on the painful region for 20 mins, followed by performing            self-care exercises to restore muscle function, stretch muscles, and coordinate muscle movement.
  - EG-post = 26 women with TMD, saliva collected post conservative treatment.
  
- Quantification of metabolites:
  - The list of 56 metabolites were processed and quantified using an NMR software, where their concentrations were       measured. These metabolite profiles were exported to excel and normalized when required.
  
- Samples:
  - The 79 samples were given specific IDs, all beginning with Amostra, followed by a capitalized letter for controls,     number beginning with zero for EG-pre, or number for EG-post.
  
#Using the data from the excel file, statistical tests were run wherein the metabolites of the various subgroups were compared to see if there are significant differences between the sub-groups for the different metabolites. Results suggested that the metabolomic profiles of women with and without TMD differ, with the EG-post group overlapping more with the control group than the EG-pre treatment. Metabolites of interest involved phenylacetate, dimethylamine, maltose, acetoin, and isovalerate, which are potential biomarkers for TMD of muscular origin.
